### 2014-06-03 TCAD2014/TCAD14.tex
TCAD14 initial commit

%
\documentclass[journal]{IEEEtran}

\usepackage{amsfonts,latexsym}

\ifx\pdfoutput\undefined
\usepackage{graphicx}
\else
\usepackage[pdftex]{graphicx}
\fi

\usepackage{epstopdf}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{multirow}

\usepackage{algorithm}
\usepackage{algpseudocode}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newtheorem{proposition}{Proposition}[section]
\newtheorem{observation}{Observation}[section]
\newtheorem{definition}{Definition}[section]

\begin{document}

\title{Title}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{ }

%\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell is with the Department of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta, GA, 30332 USA.}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2012}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}

\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}

\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
This paper is collocated into the reconfigurable computing context. An important problem into the reconfigurable computing is to analyse the error propagation when the circuit ``under test'' is setted with values of the frequency too high or in disagree with the values of the delay. The paper investigates the propagation of the error caused by a timing violation into a digital online multiplier. \\

...................\\

This work starts from the results obtained in a previus work on the Radix-2 Digit-parallel Online Multiplier (see \cite{onlinearith}).\\
 In \cite{onlinearith} the autors have provided methodologies to implement a parallel version of the online multiplier and they have proposed a probabilistic model describing the propagations of the error caused by timing violation conditions. They also backed up the models with experimental results from an image processing application demostrating that this novel design methodology can lead to substantial performance benefits compared to the design method using conventional arithmetic. The purpose of this paper is to provide a numerical model estimating the error on the online multiplier and a model to estimate the probability of timing violation in order to provide an ``a priori'' estimation for the error expectation for different values of the circuit parameters (frequency, number of digits, online delay, etc).  \\
In this paper we provide a formal model of the Overclocking Error for the radix-2 digit-parallel Online Multiplier. We also proved a mathematical '``support'' to the experimental results obtained in  \cite{onlinearith} and, in particular, we estimate the expectation of the overclocking error for these results. We provide a model describing the probablity of timing violation as ``chain start probability'' and ``chain rule probability'' representing the probability that a chain starts during the operation and the probability that a chain starts $\&$ perpetuals for a fixed number of digits, respectively.  \\

\section{Background}\label{Sec:Background}

Error models............(probably only \cite{onlinearith})..................

\subsection{Radix-2 Digit-parallel Online Multiplier}

Originally online arithmetic was designed for digit-serial operation. In order to generate the first output digit, $\delta$ digits of inputs are required. $\delta$ is known as the online delay, which is a constant, independent of the precision in a given operation. For ease of discussion, the input to our circuit is normalized to a fixed point number in the range $(-1,1)$. Based on this premise, the online representations of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}) for $j\in[-\delta,N-1]$ and $r$ denotes the radix.

\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD-first operation in online arithmetic is possible because they adopt a redundant number system, in which each digit is represented with a redundant digit set $\{-a, \cdots,-1, 0, 1, \cdots, a\}$ where $a\in[r/2,r-1]$. Due to the redundancy, the MSDs of the result can be estimated only based on partial information of the inputs and then the value of the number can be revised by the following digits.

Typically the online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm \ref{Algorithm:OnlineMult} where both inputs and outputs are $N$-digit number as represented in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case we have $\delta=3$, and $sel()$ is given by (\ref{Eq:SelFunc_OM}), from which it can be seen that the decision is made only based on 1 integer bit and 1 fractional bit of $W_{[j]}$.
\begin{equation}\label{Eq:SelFunc_OM}
\begin{array}{ll}
  sel(W_{[j]})=
    1 & if \ W_{[j]} \geq \frac{1}{2} \\
    0 &  if  \ -\frac{1}{2}\leq W_{[j]}<\frac{1}{2} \\
    \overline{1} & if \ W_{[j]}<-\frac{1}{2}
\end{array}
\end{equation}

\vspace{0.5cm}

\begin{algorithm}{\bf Online Multiplication} \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
 \State  {\bf Initialization:}   $X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$
  \State  {\bf Recurrence:}   $for~~ j=-\delta,~-\delta+1,~\cdots,~N-1 ~~do$
 \State  {\bf Compute:}    \label{Eq:OnlineMult_General}
\State  $    H_{[j]}  = r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\label{Eq:OnlineMult_H}
      \State   $  W_{[j]}    = P_{[j]} + H_{[j]}$
     \State    $  z_j        =  sel(W_{[j]})$
    \State     $  P_{[j+1]}  =  r\left(W_{[j]}-z_j\right)$\label{Eq:OnlineMult_P}
  \end{algorithmic}
\end{algorithm}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=7cm,height=4.5cm]{./Figures/circuit}
\caption{  . . . . . . . . . . . . }\label{circuit}
\end{center}
\end{figure}

..........................................\\

Let $\mu_{OM}$ denote the worst-case delay of an OM. It follows that if the clock period $T_S$ is greater than $\mu_{OM}$, correct results will be sampled. If, however, faster-than-rated sampling frequencies are applied such that $T_S<\mu_{OM}$, timing violations might happen and intermediate results will be sampled, potentially generating errors. For a given $T_S$, the maximum length of error-free propagation is described by (\ref{Eq:MaxChainLength}) where $f_S$ denotes the sampling frequency. We always ensure that the first digit of the product will be generated correctly, i.e. $b>\delta$.
%
\begin{eqnarray}\label{Eq:MaxChainLength}
\footnotesize
  b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
  %b:=\left\lceil T_S/{\mu} \right\rceil=\left\lceil {1}/{(\mu\cdot f_S)}\right\rceil
\normalsize
\end{eqnarray}
%
We now determine when this timing constraint is not met and the size of error in this case.\\

\section{Probability of Timing Violations}\label{section:Prob}

For an $N$-digit OM, let a propagation chain be generated at stage $S_{\tau}$ with the length of $d(\tau)$ digits, then the stage number $\tau$ is bounded by (\ref{Eq:Bound_GenChain}). The presence of timing violation requires $d(\tau)>b$. Besides, the chain cannot propagate over $S_{N-1}$. Thus the bound of parameter $d(\tau)$ is given by (\ref{Eq:Bound_ChainLength}).

\begin{equation}\label{Eq:Bound_GenChain}
 -\delta\leq \tau\leq N-1-b=\tau_{max}
\end{equation}

\begin{equation}\label{Eq:Bound_ChainLength}
  b<d(\tau)\leq N-1-\tau
\end{equation}
%

\noindent However, the actual length of a ``carry'' chain is dependent upon input patterns. We then examine the relationship between $d(\tau)$ and the inputs that corresponds to the generation, propagation and annihilation of this carry chain.  Let the specific input pattern of stage $S_{\tau}$ be represented by $C(\tau)$, which can be classified into four types, as listed in (\ref{Eq:class}).

\begin{equation}\label{Eq:class}
\footnotesize
    C(\tau)= \left\{ \begin{array}{llll}
        Case 1 & (C_1(\tau)): & x_{\tau +\delta+1}=0, & y_{\tau +\delta+1}=0\\
        Case 2 & (C_2(\tau)): & x_{\tau +\delta+1}\neq 0, & y_{\tau +\delta+1}\neq 0\\
        Case 3 & (C_3(\tau)): & x_{\tau +\delta+1}\neq 0, & y_{\tau +\delta+1}=0\\
        Case 4 & (C_4(\tau)): & x_{\tau +\delta+1}=0, & y_{\tau +\delta+1}\neq 0
\end{array} \right.
\normalsize
\end{equation}

\noindent The classification is based on whether a digit is zero, as all internal signals are reset to zero initially. Under the assumption that all digits are mutually independent and uniformly sampled from the digit set  $\{\overline{1},0,1\}$ as given in Section~\ref{Sec:Background}, the probabilities of $C_1,...,C_4$ are given in (\ref{Eq:probcaser}).

\begin{equation}\label{Eq:probcaser}
\begin{array}{ll}
C_1(\tau): \frac{1}{(2a+1)^2}, &  C_2(\tau): \frac{(2a+1)^2-2(a+1)-1}{(2a+1)^2}\\
C_3(\tau):\frac{a+1}{(2a+1)^2}, & C_4(\tau):  \frac{a+1}{(2a+1)^2}
\end{array}
\end{equation}

Notice that no carry chain will be generated under Case~1, because $P_{[\tau+1]}=0$. In other words, a chain could be generated if one of  the cases $C_2(\tau)$,...,$C_4(\tau)$ occur. In this case we provide the following propositions which describe the probabilities of timing violations.

\begin{proposition} (Chain rule probability)\\
Let $Pr_{\tau,d(\tau )}$ denote the probability that a carry chain start at stage $S_{\tau}$, and it have lenght $d(\tau )$, $\forall \tau \in [-\delta , N-1-b]$, $\forall d(\tau) \in [b , N-1-\tau]$, then it is given by (\ref{Eq:PChainRule})

\begin{equation}\label{Eq:PChainRule}
\footnotesize
Pr_{\tau,d(\tau )} =\left(\frac{1}{(2a+1)^2}\right)^\tau \left(\frac{(2a+1)^2-1}{(2a+1)^2}\right)^{d(\tau )}
\normalsize
\end{equation}

\noindent Proof:\\

Let $E_2$ be the event ``the carry chain is generated at stage $S_{\tau}$ and it have length $d(\tau )$'', we have

\begin{equation}
Pr_{\tau,d(\tau )} = Pr(E_2)=  \underbrace{Pr(C_1)\cdots  Pr(C_1)}_{\tau -1 \ times}\cdot
\end{equation}
$$
\cdot  \underbrace{ Pr(C_2 \ or \ C_3   \ or \ C_4)\cdots Pr(C_2  \ or \   C_3   \ or \   C_4)}_{d(\tau) \  times}   Pr(C_1)$$

that is

\begin{equation}\label{Eq:PChainRule_Exp}
\scriptsize
Pr_{\tau,d(\tau )} = Pr(C_1)^{\tau -1} Pr(C_2 \cup C_3 \cup C_4)^{d(\tau )} Pr(C_1)
\normalsize
\end{equation}

Substituting (\ref{Eq:probcaser}) into \ref{Eq:PChainRule_Exp} yields (\ref{Eq:PChainRule_raw}).

\begin{equation}\label{Eq:PChainRule_raw}
\scriptsize
Pr_{\tau,d(\tau )}= \left(\frac{1}{(2a+1)^2}\right)^{\tau -1} \left(\frac{(2a+1)^2-1}{(2a+1)^2}\right)^{d(\tau )} \frac{1}{(2a+1)^2}
\normalsize
\end{equation}

Simplifying (\ref{Eq:PChainRule_raw}) will give (\ref{Eq:PChainRule}).

\hspace{7cm} $\clubsuit$
\vspace{1ex}

\end{proposition}

For all possible values of $\tau$ and $d(\tau)$, the possibility that timing violaions happen is given by $Pr$ in (\ref{Eq:Prob_timingViol}). According to (\ref{Eq:Bound_GenChain}) and (\ref{Eq:Bound_ChainLength}), the ranges in which the parameters $\tau$ and $d(\tau)$ are defined depend upon the value of the swamping period $T_S$. Hence $Pr$ is a function of $T_S$.

\begin{equation}\label{Eq:Prob_timingViol}
Pr= \sum_\tau \sum_{d(\tau)} Pr_{\tau,d(\tau )}
\end{equation}

\section{Error Estimation}

In the presence of timing violation, multiple chains might not be correctly propagated in the online multiplication, resulting in overclocking errors generated from LSDs. Let $z_i$ and ${z_i}'$ denote the correct value and the actual value of the output digit at the stage $S_i$, respectively. Then we have ${z_i}'=z_i+\varepsilon_i$ where $\varepsilon_i$ is referred to as the overclocking error. We now locate the first output digit $z_{\lambda}$ that contains error. For a given $T_S$ and a minimum value of $\tau$ such that $d(\tau)_{max}>b$, we have $\lambda=\tau+d(\tau)_{max}+1$ for chain annihilation.
%
%. Let us consider the first chain that timing violation may occur, i.e. $d(\tau)_{max}>b$, denote that it annihilates at $S_{\lambda}$ where $\lambda=\tau+d_{max}(\tau)-1$. Therefore overclocking errors may happen from digit $\lambda$ to digit $N-1$ of the result.
%
As such, the overclocking error can be expressed by (\ref{Eq:ErrorMagnitude})

%, where $z_i$ and ${z_i}'$ denote the correct value and the actual value of the output digit of $S_i$, separately, and $\varepsilon_i\in\{\pm2^{-i},\pm2^{-i+1}\}$ since $z$ is represented using digit set $\{\overline{1},0,1\}$.
%
\begin{eqnarray}\label{Eq:ErrorMagnitude}
\scriptsize
 \left| \varepsilon\right|= \left|\sum_{j=\lambda}^{N} r^{-j}\varepsilon_j\right|
\normalsize
\end{eqnarray}

We would like to investigate the statistical characteristics of the $\varepsilon$. In general, the distribution of $\varepsilon$ can be obtained from the convolution of variable $\varepsilon_j$ where $j\in[\lambda,N]$. Each variable $\varepsilon_j$ is sum of uniform iid variables with mean zero, therefore its mean is zero and the mean of $\varepsilon$ is zero as well.

Let $v\varepsilon$ and $v\varepsilon_j$ denote the variances of variable $\varepsilon$ and $\varepsilon_j$, respectively. We can form the relationship between them as given in (\ref{Eq:VarianceErrorMagnitude}) based on (\ref{Eq:ErrorMagnitude}).

\begin{eqnarray}\label{Eq:VarianceErrorMagnitude}
\scriptsize
	v\varepsilon=\sum_{j=\lambda}^{N} r^{-2j}v\varepsilon_j
\normalsize
\end{eqnarray}

Normally for a given algorithm, in this case the Algorithm~\ref{Algorithm:OnlineMult}, the value of $\varepsilon_j$ can be determined based on the inputs, which are uniformly distributed with the digit set $\{\overline{a},\cdots,0,\cdots,a\}$. Hence for a single digit of inputs $X_{[j]}$ and $Y_{[j]}$, its variance can be denoted by $v\varepsilon^{in}$, which is a fixed value with respect to a given value of $a$. Let the variances of the input signals $X_{[j]}$ and $Y_{[j]}$ be denoted by $v\varepsilon^y_j $ and $ v\varepsilon^x_j$, respectively. Then we have (\ref{Eq:var_XY}) according to (\ref{Eq:Online_Operands}).

\begin{equation}\label{Eq:var_XY}
{v\varepsilon}^y_j = {v\varepsilon}^x_j =\sum_{i=1}^{j+\delta} r^{-2i} v\varepsilon^{in}
\end{equation}

In Algorithm~\ref{Algorithm:OnlineMult}, the variances of the input data propagate through for all iterations. For a single iteration $j$, we can calculate $v\lambda_j$ based on $v\varepsilon^y_j $ and $ v\varepsilon^x_j$. For example initially the variance of $H_{[j]}$ is given by (\ref{Eq:step4var}). 
%which is then used to generate $W_{[j]}$ with the variance given in ().

\begin{equation}\label{Eq:step4var}
{v\varepsilon}^H_j =r^{-2\delta}(v\varepsilon^{in}  v\varepsilon^y_{j+1} +v\varepsilon^{in}  v\varepsilon^x_j) 
\end{equation}

This value propagates throughout the entire iteration and finally we have the value of $v\varepsilon_j$ in (\ref{Eq:varerror}).

\begin{equation}\label{Eq:varerror}
v\varepsilon_j=r^2 \left( r^{-2\delta}\cdot {v\varepsilon}^H_j \right)
\end{equation}

Combining (\ref{Eq:step4var}) and (\ref{Eq:varerror}) yields

\begin{equation}\label{Eq:varerror_final}
v\varepsilon_j=r^2 \left( r^{-2\delta} \left( v\varepsilon^{in}  v\varepsilon^y_{j+1} +v\varepsilon^{in}  v\varepsilon^x_j \right) \right).
\end{equation}

Finally, (\ref{Eq:varerror_final}) can be used to derive the expression of $v\varepsilon$ in (\ref{Eq:varfinal}) according to (\ref{Eq:VarianceErrorMagnitude}). Notice that this expression forms the relationship between the variance of overclocking error and the given input distribution.

\begin{equation}\label{Eq:varfinal}
\begin{array}{l}
v\varepsilon =  \mathcal{F}(T_S, \mu, v\varepsilon^{in}, \delta)\\
=\sum_{k=\lambda}^{N}r^{-2k} \left( r^{2(1-\delta)} \left(v\varepsilon^{in}    v\varepsilon^y_{k+1} +v\varepsilon^{in}   v\varepsilon^x_k\right)\right)
\end{array}
\end{equation}

\section{Model Verification with FPGA results}

In this Section, we compare the proposed models for both the probability and the magnitude of overclocking errors with the experimental results from FPGAs. In the verification, we choose the redundant digit set $\{\overline{1},0,1\}$, which is most commonly used for binary redundant representations.

\subsection{Verification of Models for Error Probability}
We initially verify the models for the probability of timing violations. With the usage of the redundant digit set $\{\overline{1},0,1\}$, we have $a=1$. Substituting this in (\ref{Eq:PChainRule}) yields (\ref{Eq:OMerrorProb}).

\begin{equation}\label{Eq:OMerrorProb}
 \quad Pr_{\tau,d(\tau )} =\left(\frac{1}{9}\right)^\tau \left(\frac{8}{9}\right)^{d(\tau )} 
\end{equation}

Combining (\ref{Eq:OMerrorProb}) and (\ref{Eq:Prob_timingViol}) we obtain the values of the probability of timing violations with respect to a variety of sampling period $T_S$. The results of an 8-digit OM are shown in Fig.~\ref{Fig:ProbTimingViol}. For comparison, we also plot the results obtained from a Virtex-6 FPGA, and the data are obtained from post place-and-route simulations. It can be seen that the modelled probability values match well with the experimental results.

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=9cm,height=4.5cm]{./Figures/probTimingViolationFPGA.eps}
\caption{ Probability of timing violation for $N=8$ and $\delta=3$ }
\label{Fig:ProbTimingViol}
\end{center}
\end{figure}

\subsection{Verification of Models for Error Magnitude}
We then verify the proposed models for the estimation of timing errors. Fig.~\ref{FigerrorC} shows the distribution of the error $\varepsilon$ on the FPGA results of the Algorithm 1 with input data uniformly distributed into the digit set $\{\overline{1},0,1\}$. \\

\begin{figure}[tbp]
	\centering
	\includegraphics[width=8cm,height=5cm]{./Figures/FigerrorC}
\caption{Distribution of overclocking errors with respect to a variety of sampling period. The Xlable should be ``Sampling period'', which varies from 2ns to 3.5ns, and the Ylabel is ``Error Values''. }
	\label{FigerrorC}

\end{figure}

Several slice through Fig.~\ref{FigerrorC} are presented in Fig.~\ref{Fig:gaussianOnlinedistr} for different sampling period values and the corresponding variances of overclocking errors. We observe that the profile of the error distribution changes as the value of the sampling period changes, and the width of the bell changes as well also (see Fig.~\ref{Fig:gaussianOnlinedistr}). The value of the mean is zero for all the values of the sampling period, while the variance of the distribution changes as the sampling period changes. In particular, the values of the variance decrease as the sampling period increases in agreement with the information provided by the values of the probability of timing violation.\\

\begin{figure}[tbp]
\centering
	\includegraphics[width=9cm,height=6cm]{./Figures/gaussianOnlinedistr}
	\caption{Several slice through Fig.~\ref{FigerrorC} that shows error distribution with respect to a variety of sampling periods and the corresponding variances of overclocking error. The results are obtained from experiments.}
	\label{Fig:gaussianOnlinedistr}
\end{figure}

We have compared these experimental results with the results provided by the model defined in (\ref{Eq:varfinal}). The Figure \ref{Error_Function_Spline_experiment} shows the curve provided by the model (blue line) and the experimental results (red stars). It can be seen that the modelled results are close to the experimental results.

%the average of the values of the error obtained by comparing the two different values for each value of the sampling period is the $O(10^{-3})$. \\

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=8cm,height=5cm]{./Figures/Error_Function_Spline_experiment}
\caption{Variances of overclocking errors. The results are obtained from the proposed models (blue line) and the experiments (red stars). (Legend and xylabels to be added.) }\label{Error_Function_Spline_experiment}
\end{center}
\end{figure}

%\noindent The {\bf Expectation of Overclocking Error} is also function of the sampling period and it is obtained by combining the values of $Pr$ and $\varepsilon$ as in (\ref{eovc}).

%\begin{equation}\label{eovc}
%E_{ovc}(T_S)= Pr(T_S) \cdot  v(T_S)
%\end{equation}

%\section*{Acknowledgment}

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{onlinearith} Datapath synthesis for overclocking: Online Arithmetic for Latency-Accuracy Trade-offs.
\bibitem{convol} J. B. Uspensky, Introduction to Mathematical Probability (New York: McGraw-Hill, 1937)
\bibitem{ross} Ross, Introduction to Probability Models, University of Southern California - Los Angeles (CA), 2010 ELSEVIER
\bibitem{spline} C. de Boor, A Practical Guide to Splines, Springer-Verlag, 1978.
\bibitem{kendall} Kendall E. Atkinson, On the order of convergence of natural cubic spline interpolation, SIAM Journal on Numerical Analysis, Vol 5 No 1, 1968

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

\end{document}